Parallel Information Set Generation for Kriegspiel

### 2012-07-16 sbac/paper.tex
minor

%\documentclass[times, 10pt,twocolumn]{article}
%\documentclass[conference]{IEEEtran}
\documentclass[10pt, conference, compsocconf]{IEEEtran} 
%\usepackage{latex8}
%\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage[tableposition=top,font=bf]{caption}

@@

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{schofield12hyperplay}.

@@

For many specific games, solving the information set generation problem is
trivial.  For example, in a Poker game, the unseen cards held by a player's
opponents may be any permutation of the cards not seen by that player (i.e.,
hole cards and any revealed community cards).  After the betting rounds, it
would not be reasonable to assume that each of these permutations of unseen
cards is equally probable, as betting decisions made by the players up to that
point would be affected by the quality of those players' cards.  But the set of
{\em possible} hands for all of the opponents is easy to conceive and
enumerate.

@@

The remainder of the paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
necessary background material, including the rules of Kriegspiel. Next, in
Section~\ref{info}, we describe the nature of the information set generation
problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our
parallel implementation, including the {\sc Charm++} Parallel State Space
Search Engine (ParSSSE) platform.  We then provide our experimental results in
Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We
conclude with a discussion of future work.

@@

\begin{figure}
\begin{center}
\begin{verbatim}
1. d4 {(:)}
   a5 {(:)}
2. Bg5 {(:)}
   b6 {(:)}
3. Nc3 {(:Bd8)}
   c5 {(:)}
4. d5 {(:Bd8;Tc5)}
   Na6 {(:)}
5. d6 {(:)}
   f6 {(:Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   exd6 {(Xd6:Td6)}
8. Qd5 {(:)}
   Qc7 {(:)}
9. Qf7+ {(:)} 
   Kd8 {:g6,Kxf7,Ke7}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\end{center}
\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves. The final, legal move for each player is shown followed by the list of moderator declarations and all moves that were attempted but disallowed before the legal move was accepted.}
\label{listing}
\end{figure}

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\end{figure}

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Algorithm}
\label{codelisting1}
\end{figure}

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

Some chares may inherently require more computation than others, but the amount
of computation associated with each piece of work is not always known before
execution.  In tree search, for example, a piece of work may be defined as the
computation of a minimax value on a node in the tree using alpha-beta
pruning~\cite{knuth75analysis,baudet78analysis}.  A natural decomposition is to
assign different nodes to different processors for evaluation.  Because some
nodes will be pruned at shallower levels of the tree than others, the amount of
work performed at each subtree can vary considerably.  In a traditional
parallel framework, this would result in many processors idling while a small
number of processors execute the larger workloads. By allowing the {\sc
Charm++} runtime system to schedule a large number of small pieces of work,
better adaptive load balancing is achieved.

@@

A {\em grainsize control} parameter specifies a bound on the size of a piece of
work that can be passed along to the runtime system for scheduling (in other words, computational granularity of a task which will be performed sequentially).  Once a
task has been subdivided into an appropriate number of subtasks, those tasks
run to completion on a single processor without preemption and without generating
any additional parallel tasks.  If the grainsize is too large, then the number
of pieces of work to be done may be lower than the number of processors
available, forcing some processors to idle.  If the grainsize is too small,
then the runtime system is flooded with a large number of trivial tasks, and
application performance is degraded because of the overhead of creating and scheduling tasks
instead of actually doing them.  {\sc ParSSSE} supports adaptive grainsize
management, so that the threshold for sequential processing can be changed
during the execution of the program, depending on the properties of the problem
and/or the realtime performance analysis by the runtime system.

@@

The advantage of the work stealing strategy is that the message-passing
overhead is only incurred if at least one processor is idling.  Furthermore,
because victims respond by sending the highest priority work, the most
important tasks are scheduled sooner.

@@

\begin{table}[ht]

@@

In our next round of experiments, we evaluated the impact of load-balancing
strategies on our application's performance.  We used the random, work stealing, and
neighborhood balancing strategies that are available in the general purpose
search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
show that there was actually little variation in performance across the
different strategies.  Across all of our experiments, the work stealing
approach appeared to be consistently --- if only marginally --- better than the
others.  Our hypothesis here is that the nodes are rarely starved for work.  In
other words, perhaps the very nature of the problem means that the load is
naturally well-balanced and that the load-balancer rarely needs to do a
significant amount of redistribution.  This is also related to the
fact that we are doing an all-solutions search. 
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}

\caption{Effect of load balancing strategy}
\label{3schemes}
\end{figure}

\subsubsection{Effect of Grainsize}
\label{grainsize}

\begin{figure*}[ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\caption{Grainsize Control for different problem sizes. Sequential threshold denoted the depth in the search tree after which the remainder of the task (game tree search) is performed sequentially. Grain size = max game tree depth - sequential threshold.}
%\vspace{-0.1in}
\label{fig:dep}
\end{figure*}
%

@@

@@

We next study the variation in the efficiency of the parallelization of the
algorithm as size of the problem increases.  Here, we varied the number
of plies in the test instance from 8 to 12. We used the work stealing
load-balancing strategy.

The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
there is not enough additional parallel work to justify the jump from 256 to
1,024 processors, and performance actually degrades.  Aside from this anomaly,
however, we see consistent speedups as we increase the number of processors. 
The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem. Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency. As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom). As the number of processors increases, parallel efficiency decreases for constant problem size. Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.
\begin{figure}[ht] 
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthspe.pdf}
\end{minipage}
\caption{(Top) Execution time for three different sizes of problem as the number of cores used increases.
(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency}
\label{3depths}
\end{figure}
\subsection{Real Game Instance}
%\begin{figure}[ht]
% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
%
%\caption{Kriegspiel position from Virgil-David 1969 with information set size
%764,209.}
%
%\label{prob4} 
%\vspace{-0.2in}
%\end{figure}
%
%\begin{figure}[ht]
% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
%
%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
%Chess championship.  Information set size is 5,099,257.}
%
%\label{prob5} 
%\vspace{-0.2in}
%\end{figure}

We also evaluated our information set generation routine on 40 positions from
actual Kriegspiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth.

These results confirm the observations made in Section~\ref{grainsize}. The optimal sequential threshold varies for the different numbers of cores and
the pattern is consistent across problem instances.  The optimal
sequential threshold increases as the number of cores increases.  This is expected,
because as this parameter grows, the number of distinct pieces of work
increases.  When the number of chares per processor is high, much of
the overhead of parallelization is wasted.  But as the number of cores
increases, the work can be efficiently broken up into smaller pieces and spread to all the cores.  With more pieces of work, it becomes easier to balance
the load.

@@

\label{tab:prob4}
\end{table}

@@

\label{tab:prob5}
\end{table}
\begin{figure}[ht]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Performance results for Information Set Generation for 2 games of Kriegspiel}
\label{prob4} 
\end{figure}

Note that both problems scale well even up to 1,024 processors.  We achieve a
speedup of $443$ in the game between humans and up to $536$ in the game between
computers.  Figure~\ref{prob4} shows the plots of the running
times on these two problems with the best grainsize for each core setting.  These plots show that the
running time continues to decrease as the number of cores increases.

\section{Comparison with previous results, Discussion}
Our results compare favorably to previous work in parallel game tree search,
shown in Table~\ref{bestspeedups}.  The previous best parallelization results
were for parallelizing alpha-beta pruning.  Ferguson and Korf achieved a
speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.
In the game of chess, Felton and Otto showed a speedup of 100 on 256
cores~\cite{felten88highly}.  This limited scalability is due to speculative
loss.  Parts of the game tree that are searched during the parallel execution
would have been pruned had the algorithm run sequentially.  This is a
fundamental limitation in alpha-beta search.

@@

>>>>

%\documentclass[times, 10pt,twocolumn]{article}
%\documentclass[conference]{IEEEtran}
\documentclass[10pt, conference, compsocconf]{IEEEtran} 
%\usepackage{latex8}
%\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage[tableposition=top,font=bf]{caption}
%\usepackage[tableposition=top,font=bf]{caption}

@@

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
which are indistinguishable to a player since they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{schofield12hyperplay}.

@@

For many specific games, solving the information set generation problem is
trivial.  For example, in a Poker game, the unseen cards held by a player's
opponents may be any permutation of the cards not seen by that player (i.e.,
hole cards and any revealed community cards).  After the betting rounds, it
would not be reasonable to assume that each of these permutations of unseen
cards is equally probable, as betting decisions made by the players up to that
point would be affected by the quality of those players' cards.  However, the set of
{\em possible} hands for all of the opponents is easy to conceive and
enumerate.

@@

The remainder of the paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
necessary background material, including the rules of Kriegspiel. Next, in
Section~\ref{info}, we describe the nature of the information set generation
problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our
parallel implementation, including the {\sc Charm++} Parallel State Space
Search Engine (ParSSSE) platform.  We then provide our experimental results in
Section~\ref{Results} followed by comparison with previous results and some discussion in Section~\ref{discuss}.  Related work is discussed in Section~\ref{Related}.  We
conclude with a discussion of future work.

@@

\begin{figure}
\begin{center}
\begin{verbatim}
1. d4 {(:)}
   a5 {(:)}
2. Bg5 {(:)}
   b6 {(:)}
3. Nc3 {(:Bd8)}
   c5 {(:)}
4. d5 {(:Bd8;Tc5)}
   Na6 {(:)}
5. d6 {(:)}
   f6 {(:Td6)}
6. e4 {(:)}
   fxg5 {(Xg5:Td6;Tg5)}
7. Bc4 {(:)}
   exd6 {(Xd6:Td6)}
8. Qd5 {(:)}
   Qc7 {(:)}
9. Qf7+ {(:)} 
   Kd8 {:g6,Kxf7,Ke7}
10. Qxf8++ {(1-0:)}
\end{verbatim}
\end{center}
\caption{Listing of a Kriegspiel game resulting in checkmate by white after 10 moves. The final, legal move for each player is shown followed by the list of moderator declarations and all moves that were attempted but disallowed before the legal move was accepted.}
\label{listing}
\vspace{-0.1in}
\end{figure}

%\SubSection{Example}
\begin{figure}[ht]
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/9B.png}
\caption{Actual Position}
\label{fig:figure1}
\end{minipage}
\hspace{0.5cm}
\centering
\begin{minipage}[b]{0.45\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{images/OneView.png}
\caption{White's view}
\label{fig:figure2}
\end{minipage}
\vspace{-0.2in}
\end{figure}

@@

\textbf{The Information Set Generation}
(ISG) operation can be formulated as a combinatorial search problem.  The input to the algorithm is a list
of $t$ observations for one player $k$ (denoted $o_{1:t}^{(k)}$ in a Kriegspiel game.    
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\end{algorithmic}
\end{boxedminipage}
\caption{Core of Information Set Generation Algorithm}
\label{codelisting1}
\end{figure}

@@

\section{Parallel Implementation using Charm++ and ParSSSE}\label{ParSSSE}
We built our parallel ISG search engine on top of the {\sc Charm++} run-time
system for parallel processing~\cite{kale93charm,kale09charm}.  {\sc Charm++}
is an extension to the C++ programming language that provides
machine-independent infrastructure for parallel computation.  The language and
its accompanying runtime system have been ported to many shared-memory and
distributed-memory platforms.

@@

Some chares may inherently require more computation than others, however the amount
of computation associated with each piece of work is not always known before
execution.  In tree search, for example, a piece of work may be defined as the
computation of a minimax value on a node in the tree using alpha-beta
pruning~\cite{knuth75analysis,baudet78analysis}.  A natural decomposition is to
assign different nodes to different processors for evaluation.  Since some
nodes will be pruned at shallower levels of the tree than others, the amount of
work performed at each subtree can vary considerably.  In a traditional
parallel framework, this would result in many processors idling while a small
number of processors execute the larger workloads. By allowing the {\sc
Charm++} runtime system to schedule a large number of small pieces of work,
better adaptive load balancing is achieved.

@@

A {\em grainsize control} parameter specifies a bound on the size of a piece of
work that can be passed along to the runtime system for scheduling (in other words, computational granularity of a task which will be performed sequentially).  Once a
task has been subdivided into an appropriate number of subtasks, those tasks
run to completion on a single processor without preemption and without generating
any additional parallel tasks.  If the grainsize is too large, then the number
of pieces of work to be done may be lower than the number of processors
available, forcing some processors to idle.  If the grainsize is too small,
then the runtime system is flooded with a large number of trivial tasks, and
application performance is degraded due to the overhead of creating and scheduling tasks
instead of actually doing them.  {\sc ParSSSE} supports adaptive grainsize
management, so that the threshold for sequential processing can be changed
during the execution of the program, depending on the properties of the problem
and/or the realtime performance analysis by the runtime system.

@@

The advantage of the work stealing strategy is that the message-passing
overhead is only incurred if at least one processor is idling.  Moreover,
since victims respond by sending the highest priority work, the most
important tasks are scheduled sooner.

@@

@@

In our next round of experiments, we evaluated the impact of load-balancing
strategies on our application's performance.  We used the random, work stealing, and
neighborhood balancing strategies that are available in the general purpose
search engine.  These results are shown in Figure~\ref{3schemes}.  Our results
show that there was actually little variation in performance across the
different strategies.  Across all of our experiments, the work stealing
approach appeared to be consistently --- if only marginally --- better than the
others.  Our hypothesis here is that the nodes are rarely starved for work.  In
other words, perhaps the very nature of the problem means that the load is
naturally well-balanced and that the load-balancer rarely needs to do a
significant amount of redistribution.  This is also related to the
fact that we are doing an all-solutions search.

\subsubsection{Effect of Grainsize}

@@

\begin{table}[t]

\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}

\caption{Effect of load balancing strategy}
\label{3schemes}
\vspace{-0.2in}
\end{figure}

\label{grainsize}

\begin{figure*}[ht]
\centering
\subfigure[Depth 8]{
\includegraphics[width=2.1in]{plots/dep8.pdf}
\label{dep8}
} 
\subfigure[Depth 10]{
\includegraphics[width=2.1in]{plots/dep10.pdf}
\label{dep10}
} 
\subfigure[Depth 12]{
\includegraphics[width=2.1in]{plots/dep12.pdf}
\label{dep12}
} 
\caption{Grainsize Control for different problem sizes. Sequential threshold denotes the depth in the search tree after which the remainder of the task (game tree search) is performed sequentially. Grain size = max game tree depth - sequential threshold.}
%\vspace{-0.1in}
\label{fig:dep}
\vspace{-0.2in}
\end{figure*}
%

@@

We next study the variation in the efficiency of the parallelization of the
algorithm as size of the problem increases.  Here, we varied the number
of plies in the test instance from 8 to 12. We used the work stealing
load-balancing strategy.  
\begin{figure}[ht] 
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthspe.pdf}
\end{minipage}
\caption{(Top) Execution time for three different sizes of problem as the number of cores used increases.
(Bottom) Parallel Efficiency (wrt to 16 cores) vs number of cores, to help see the effect of problem size and iso-efficiency}
\label{3depths}
\vspace{-0.3in}
\end{figure}
The results are shown in Figure~\ref{3depths}.  In the easiest case (depth 8),
there is not enough additional parallel work to justify the jump from 256 to
1,024 processors, and performance actually degrades.  Aside from this anomaly,
however, we see consistent speedups as we increase the number of processors. 
The plot on the bottom helps to illustrate the effect of problem size on parallel efficiency (defined as Speedup/P) and helps understand the iso-efficiency of the problem. Iso-efficiency~\cite{b:grama-isoefficiency} means how should the problem size be increased with increasing number of processors to maintain same level of parallel efficiency. As the size of the problem increases, the amount of parallelizable work available also increases, and hence parallel efficiency is higher for larger problem sizes and constant number of processors (Figure~\ref{3depths} bottom). As the number of processors increases, parallel efficiency decreases for constant problem size. Figure~\ref{3depths} suggests that in parallel information set generation, to maintain constant parallel efficiency, problem size (search depth) should be roughly increased by two when the number of processor in increased by 4 times.

@@

\label{tab:prob4}
\vspace{-0.1in}
\end{table}

@@

\label{tab:prob5}
\vspace{-0.1in}
\end{table}

\subsection{Real Game Instance}
%\begin{figure}[ht]
% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
%
%\caption{Kriegspiel position from Virgil-David 1969 with information set size
%764,209.}
%
%\label{prob4} 
%\vspace{-0.2in}
%\end{figure}
%
%\begin{figure}[ht]
% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
%
%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer
%Chess championship.  Information set size is 5,099,257.}
%
%\label{prob5} 
%\vspace{-0.2in}
%\end{figure}

We also evaluated our information set generation routine on 40 positions from
actual Kriegspiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth. We used Charm++'s \texttt{net-ibverbs} machine layer build and $-O3$ optimization for these experiments.

These results confirm the observations made in Section~\ref{grainsize}. The optimal sequential threshold varies for the different numbers of cores and
the pattern is consistent across problem instances.  The optimal
sequential threshold increases as the number of cores increases.  This is expected,
because as this parameter grows, the number of distinct pieces of work
increases.  When the number of chares per processor is high, much of
the overhead of parallelization is wasted.  However, as the number of cores
increases, the work can be efficiently broken up into smaller pieces and spread to all the cores.  With more pieces of work, it becomes easier to balance
the load.

Note that both problems scale well even up to 1,024 processors.  We achieve a
speedup of $443$ in the game between humans and up to $536$ in the game between
computers.  Figure~\ref{prob4} shows the plot of the running
times on these two problems with the best grainsize for each core setting.  These plots show that the
running time continues to decrease as the number of cores increases.

\section{Comparison with previous results, Discussion}
\label{discuss}
Our results compare favorably to previous work in parallel game tree search,
shown in Table~\ref{bestspeedups}.  The previous best parallelization results
were for parallelizing alpha-beta pruning.  Ferguson and Korf achieved a
speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.
In the game of chess, Felton and Otto showed a speedup of 100 on 256
cores~\cite{felten88highly}.  This limited scalability is due to speculative
loss.  Parts of the game tree that are searched during the parallel execution
would have been pruned had the algorithm run sequentially.  This is a
fundamental limitation in alpha-beta search.

@@

\begin{figure}[t]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Performance results for Information Set Generation for two games of Kriegspiel, position ten}
\label{prob4} 
\end{figure}
